|

### 2019-09-24 analysis/paper/manuscript.Rmd
Address Keith's comments on the ms down to the Survey data section

<!-- READ AND REFERENCE: -->
<!-- - https://www.sciencedirect.com/science/article/pii/S1364815210000551 -->

<!-- AIM FOR SHORT COMMUNICATION IN Environmental Modeling & Software -->

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013]. For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way. Traditional formats (e.g. slides) of presenting information at stakeholder meetings, and to fisheries managers, are simply insufficient to convey the richness of the information available. More importantly, these static and linear formats tend to stifle meaningful discussion as important details and patterns tend to get lost under the sheer volume of output. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community. As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018]. This trend includes fisheries research and many of these tools, and the code used to produce them, are available online (see Appendix A for examples).

Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs. First, we focus on data that are commonly presented in stock assessments and use interactive plots to simplify the detailed exploration of data from long-term monitoring programs program. Second, we focus on the modeling aspect of stock assessment and demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) knowing the data, and 2) data modeling.

# Knowing the data

The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both. As such, stock assessment biologists need to manage large volumes of data from a variety of sources. Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind a plot. For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data. This is exemplified by two recently developed interactive tools: 1) a tool designed to quickly examine fishery-independent survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.

>>>>

*The greatest value of a picture is when it forces us to notice what we never expected to see.*  
- Tukey [-@tukey1977]

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013]. For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way. Traditional formats (e.g. slides) for presenting information at stakeholder meetings are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community. As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018]. This trend includes fisheries research and many of these tools, and the code used to produce them, are available online (see Appendix A for examples).

Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs. First, we focus on data that are commonly presented in stock assessments and use interactive plots to simplify the detailed exploration of data from long-term monitoring programs program. Second, we focus on the modeling aspect of stock assessment and demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling.

# Data exploration

The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both. As such, stock assessment biologists need to manage large volumes of data from a variety of sources. Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. It has long been recognized that this challenge can be mediated, to a degree, by the application of interactive visualization tools as they allow detailed exploration of the data behind a plot [@tukey1985; @fisherkeller1988]. For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data. These abilities are illustrated  by two recently developed interactive tools: 1) a tool designed to quickly examine fishery-independent survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.
